First Exclusive Measurement of Deeply Virtual Compton Scattering off $^4$He: Toward the 3D Tomography of Nuclei

### 2017-05-14 eg6-letters/coherent/dvcs_he4_prl.tex
dvcs_he4_prl.tex:  minor grammar suggestion updates (forgotten or overwritten) before resending

%
\date{\today}
\begin{abstract}
We report on the first measurement of the beam-spin asymmetry in the exclusive 
process of coherent deeply virtual Compton scattering off a nucleus. The 
experiment used the 6 GeV electron beam from the CEBAF accelerator at Jefferson 
Lab incident on a pressurized $^4$He gaseous target in front of the CEBAF large acceptance 
spectrometer (CLAS). The scattered electron was detected by CLAS and the photon 
by a dedicated electromagnetic calorimeter at forward angles. To ensure the 
exclusivity of the process, a specially designed radial time projection chamber 
was used to detect the recoiling $^4$He nuclei. We measured beam-spin 
asymmetries larger than those observed on the free proton in the same kinematic 
domain. From these, we were able to extract, in a model-independent way, the 
real and imaginary parts of the only $^4$He Compton form factor, $\cal H_A$. 
This pioneering measurement of coherent deeply virtual Compton scattering on 
the $^4$He nucleus, with a fully exclusive final state via nuclear recoil 
tagging, leads the way toward 3D imaging of the partonic structure of nuclei.
\end{abstract}
\pacs{Valid PACS appear here}

@@

The $^4$He nucleus is an ideal experimental target for nuclear DVCS, as it is 
light enough to be detected by our experimental setting, while it is subject 
to interesting nuclear effects~\cite{JSeely}. Its zero spin also leads to an 
important simplification, as a spin-zero hadron is parametrized by only one 
chiral even GPD ($H_{A}(x,\xi,t)$) at leading twist, while four GPDs arise for 
the spin-$\frac{1}{2}$ nucleon. This significantly simplifies the interpretation of the 
data and allows a model independent extraction of the $^4$He CFF 
($\mathcal{H}_{A}$) presented at the end of this letter.

The CEBAF large acceptance spectrometer (CLAS) in Hall B at Jefferson 
Lab~\cite{Mecking:2003zu} has been previously supplemented with an inner 
calorimeter (IC) and a solenoid magnet to measure DVCS observables on the nucleon
\cite{Girod:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}. The IC 
extended the photon detection acceptance of CLAS to polar angles as low as 
4$^{\circ}$. The 5~T solenoid magnet acted as a guiding field for the 
low-energy M\o{}ller electrons that were absorbed in a  
heavy shield placed around the beam line.

@@

We identified two main backgrounds, accidental coincidences and exclusive coherent
$\pi^0$ production. The accidentals have particles originating from different events,
and we estimated their contribution to be 4.1\% of our sample. We evaluated this 
contribution by selecting events passing all the cuts but with the scattered electron and 
$^4$He originating from different vertices. The $\pi^0$ production can easily
be mistaken for DVCS when one of the two photons from the $\pi^0$ 
decay is produced at low energy in the laboratory frame and remains undetected.  
To estimate the effect of this contamination, we developed an event generator 
tuned on the experimental yield of exclusive $\pi^0$ with two photons measured. 
We used this generator together with a GEANT3 simulation of our 
detectors to estimate the ratio of the number of $\pi^0$ events where the 
two photons are detected to those that are misidentified as DVCS events. This 
ratio is then multiplied by the measured yield of exclusive $\pi^0$ events to 
correct the DVCS data. Depending on the kinematics, we found contaminations of 
2 to 4\%.

@@

%conclusion
In summary, we have presented the first measurement of the beam-spin asymmetry
of exclusive coherent DVCS off $^4$He using the CLAS spectrometer supplemented 
with a RTPC. This setup allowed detection of the 
low energy $^4$He recoils in order to ensure an exclusive measurement of the
coherent DVCS process. The azimuthal dependence of the measured $A_{LU}$ has 
been used to extract, in a model independent way, the real and the imaginary 
parts of the $^4$He CFF, $\mathcal{H}_A$. The extracted CFF is in  
agreement with predictions of the available models. This first fully exclusive 
experiment opens new perspectives for studying nuclear structure with the GPD 
framework and paves the way for future measurements at JLab using 12 GeV CEBAF 
and upgraded equipment.

>>>>

%
\date{\today}
\begin{abstract}
We report on the first measurement of the beam-spin asymmetry in the exclusive 
process of coherent deeply virtual Compton scattering off a nucleus. The 
experiment used the 6 GeV electron beam from the CEBAF accelerator at Jefferson 
Lab incident on a pressurized $^4$He gaseous target in front of the CEBAF Large Acceptance 
Spectrometer (CLAS). The scattered electron was detected by CLAS and the photon 
by a dedicated electromagnetic calorimeter at forward angles. To ensure the 
exclusivity of the process, a specially designed radial time projection chamber 
was used to detect the recoiling $^4$He nuclei. We measured beam-spin 
asymmetries larger than those observed on the free proton in the same kinematic 
domain. From these, we were able to extract, in a model-independent way, the 
real and imaginary parts of the only $^4$He Compton form factor, $\cal H_A$. 
This pioneering measurement of coherent deeply virtual Compton scattering on 
the $^4$He nucleus, with a fully exclusive final state via nuclear recoil 
tagging, leads the way toward 3D imaging of the partonic structure of nuclei.
\end{abstract}
\pacs{Valid PACS appear here}

@@

The $^4$He nucleus is an ideal experimental target for nuclear DVCS, as it is 
light enough to be detected by our experimental setting, while it is subject 
to interesting nuclear effects~\cite{JSeely}. Its zero spin also leads to an 
important simplification, as a spin-zero hadron is parametrized by only one 
chiral even GPD ($H_{A}(x,\xi,t)$) at leading-twist, while four GPDs arise for 
the spin-$\frac{1}{2}$ nucleon. This significantly simplifies the interpretation of the 
data and allows a model independent extraction of the $^4$He CFF 
($\mathcal{H}_{A}$) presented at the end of this letter.

The CEBAF Large Acceptance Spectrometer (CLAS) in Hall B at Jefferson 
Lab~\cite{Mecking:2003zu} has been previously supplemented with an inner 
calorimeter (IC) and a solenoid magnet to measure DVCS observables on the nucleon
\cite{Girod:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}. The IC 
extended the photon detection acceptance of CLAS to polar angles as low as 
4$^{\circ}$. The 5-T solenoid magnet acted as a guiding field for the 
low-energy M\o{}ller electrons that were absorbed in a  
heavy shield placed around the beam line.

@@

We identified two main backgrounds, accidental coincidences and exclusive coherent
$\pi^0$ production. The accidentals have particles originating from different events,
and we estimated their contribution to be 4.1\% of our sample. We evaluated this 
contribution by selecting events passing all the cuts but with the scattered electron and 
$^4$He originating from different vertices. The $\pi^0$ production can easily
be mistaken for DVCS when one of the two photons from the $\pi^0$ 
decay is produced at low-energy in the laboratory frame and remains undetected.  
To estimate the effect of this contamination, we developed an event generator 
tuned on the experimental yield of exclusive $\pi^0$ with two photons measured. 
We used this generator together with a GEANT3 simulation of our 
detectors to estimate the ratio of the number of $\pi^0$ events where the 
two photons are detected to those that are misidentified as DVCS events. This 
ratio is then multiplied by the measured yield of exclusive $\pi^0$ events to 
correct the DVCS data. Depending on the kinematics, we found contaminations of 
2 to 4\%.

@@

%conclusion
In summary, we have presented the first measurement of the beam-spin asymmetry
of exclusive coherent DVCS off $^4$He using the CLAS spectrometer supplemented 
with a RTPC. This setup allowed detection of the 
low-energy $^4$He recoils in order to ensure an exclusive measurement of the
coherent DVCS process. The azimuthal dependence of the measured $A_{LU}$ has 
been used to extract, in a model independent way, the real and the imaginary 
parts of the $^4$He CFF, $\mathcal{H}_A$. The extracted CFF is in  
agreement with predictions of the available models. This first fully exclusive 
experiment opens new perspectives for studying nuclear structure with the GPD 
framework and paves the way for future measurements at JLab using 12 GeV CEBAF 
and upgraded equipment.
